---
title: "An introduction to the GenomicLayers package"
author: "Dave T. Gerrard"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{An introduction to the GenomicLayers package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---



This will be an introductory vignette on how and why to use GenomicLayers.

### Introduction

Rationale...

_GenomicLayers_ simulates epi-genetic changes to DNA sequence or entire genomes. It does this by linking the DNA sequence, which cannot be changed, with one or more equally sized _layers_, which can be modified by _binding factors_. The _binding factors_ may recognise specific DNA sequences or motifs or they may recognise existing modifications already applied to the layers. After binding, the _binding factors_ may then make modifications to any or all of the layers. In this way, some regions of the sequence or genomes may change states (e.g. open vs closed) and permit or deny access to further binding factors.


### Installation

You may have already done this but if not, you can install the _GenomicLayers_ package direct from GitHub if you already have the _devtools_ package installed and working. For _devtools_ see here: ...

If you already have _devtools_, then the following commands should install _GenomicLayers_:-


```{r install-coms, eval=FALSE, echo=TRUE}
	library(devtools)
	install_github(repo="davetgerrard/GenomicLayers",build_vignettes = TRUE)
```
The above may take several minutes and requires several dependencies. If it does not work, or you are in a hurry, leave out the 'build_vignettes'. Hint: make sure _devtools_ is up to the most recent version.

```{r install-noVignettes, eval=FALSE, echo=TRUE}
	install_github(repo="davetgerrard/GenomicLayers")
```

To view the introduction vignette, type

```{r vignette-intro, eval=FALSE, echo=TRUE}
	vignette("GenomicLayersIntroduction")
```
To view what vignettes are available, type

```{r list-vignettes, eval=FALSE, echo=TRUE}
	vignette(package="GenomicLayers")
```

### Set up the target sequence or genome
Load the GenomicLayers library. 

```{r load_libs, echo=TRUE, message=FALSE}
library(GenomicLayers)    # (this command outputs a lot of information, which is hidden in this vignette)

```

_GenomicLayers_ is designed to work on whole eukaryotic genomes including human and mouse but as these are rather large, for the purpose of demonstration we will use a few chromosomes from the _Saccharomyces cerevisiae_ (Yeast) SacCer3.

```{r load-yeast, echo=TRUE}
library(BSgenome.Scerevisiae.UCSC.sacCer3)
genome <- BSgenome.Scerevisiae.UCSC.sacCer3
seqinfo(genome)

sequences_to_keep <- c("chrI", "chrII", "chrIII")   
# chrIII also contains the mating type loci HMR and HML
genomeSub <- keepBSgenomeSequences(genome, sequences_to_keep)
genomeSub    # this should now still be a useable BSgenome object but with only three  chromosomes.  


```


### Create some layers on our genome

This is where we use the sequence and build some layers on it. 

```{r new_layers, echo=TRUE}

sacCer.LS <- createLayerSet.BSgenome(genomeSub,
                                     n.layer=2, 
                                     layer.names=c("recruiter", "promoter"))
```

### Create some binding factors  

This is where we specify some binding factors that can bind to the sequence and, optionally, modify the layers (but not the sequence).

If you want to learn about pattern matching against DNA sequences, we recommend that you check out the vignettes in the Biostrings package. Perhaps also have a look at PWMenrich, MOTIV etc.

We first create a binding factor that can recognise a DNA motif for a TATA box, we'll call it bf1.tata and then test it against the layerSet object we've already created. It should create an GRanges or hits object with the locations for all matches to the sequence TATAA. We'll also specify that this binding factor should mark those positions on the layer _recruiter_. "TATAWAWA"

```{r new_bf1}
bf1.tata <- createBindingFactor.DNA_consensus(name="bf1.tata", patternString = "TATAWAWA", fixed =FALSE,  mod.layers="recruiter",mod.marks=1) 
#print(bf1)
matchBindingFactor.BSgenome(layerSet=sacCer.LS, bindingFactor=bf1.tata)
```

Then create a second binding factor that does not recognise a particular sequence, but which can bind to marks on the _recruiter_ layer . If we test this binding factor alone, it should produce a hits object with no hits because the _recruiter_ layer starts with not marked regions.

```{r new_bf2}
bf2.tataBox <- createBindingFactor.layer_region(name="bf2.tataBox", 
                          profile.layers="recruiter",profile.marks=1,                                                   mod.layers="promoter",mod.marks=1,  patternLength = 5) 
#print(bf2)
matchBindingFactor.BSgenome(layerSet=sacCer.LS, bindingFactor=bf2.tataBox)
```

To use multiple binding factors together in a sequence, we add them into a named list. IMPORTANT: currently, this list must have names, so typically we specify the names when building the list. The list names do not have to agree with the _$name_ property of the individual binding factors (this allows the same binding factor to be re-used without re-specifying the model).

Additionally, create a vector of integer abundances that determine the effective activity of each binding factor. This can be fixed or varied throughout the simulation.
```{r bf_list}
bf_list <- list(bf1.tata=bf1.tata, bf2.tataBox=bf2.tataBox)
#print.bfSet(list(bf1))  # not working in package. FIX this .

bf_abund <- structure(c(500, 500), names=names(bf_list))
```

### Run a sequence of layer binding

Run a simulation and collect the output

```{r run_lb}
sim.sacCer.LS <- runLayerBinding.BSgenome(layerList=sacCer.LS, factorSet=bf_list,
                                          bf.abundances = bf_abund, verbose=TRUE)

```

Hopefully, the layer called 'promoter' should now be marked in some places.

```{r inspect_lb}

sim.sacCer.LS$layerSet[['promoter']]

```

### Visualise the results

What happened in the simulation?  

Are the marked regions anywhere near the real Yeast promoters?  - No.


### Using larger genomes

Get to know the BSgenome packages....

