---
title: "MouseX-inacivation"
author: "Dave T. Gerrard"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


## Introduction


A real model of spreading of several factors across the X chromosome.


Requires mouse BSgenome.Mmusculus.UCSC.mm9 which can be obtained from Bioconductor

  source("https://bioconductor.org/biocLite.R")
  biocLite("BSgenome.Mmusculus.UCSC.mm9")


## Load prerequisites


If  _BSgenome.Mmusculus.UCSC.mm9_ and _GenomicLayers_ are installed, the following should not give errors.

```{r load-libs, echo=TRUE}
library(BSgenome.Mmusculus.UCSC.mm9)
library(GenomicLayers)
```




## Set up chromosome and layers
```{r }
#OUTPUTDIR <- "GL_out/results/mus_X_inactivation/"
GENOME <- 'mm9'

# load X chromosome sequence 
genome <- BSgenome.Mmusculus.UCSC.mm9 


# set up a layerSet on chrom X.
thisChrom <- genome[["chrX"]] 
layerList.X <- createLayerList.DNAstring(seq=thisChrom, 
                                layerNames=c("CpG_island","PRC","H3K27me3"))

```


## Load/Create binding factors
```{r }
#N.B. temporarily setting a fixed patternLength, 
# because don't know how to implement variable patternLength yet.
# current effect is that hits < patternLength are discarded
bf.CpGisland <- createBindingFactor.DNA_regexp("CpGisland", patternString="(CG.{0,20}){9}CG",
                                               patternLength=20, mod.layers = "CpG_island", 
                                               mod.marks=1, stateWidth=200)

bf.PRC <- createBindingFactor.layer_region("PRC",  patternLength=200, mismatch.rate=0, 
                                           profile.layers = "CpG_island", profile.marks = 1,  
                                           mod.layers = "PRC", mod.marks=1, stateWidth=500)

# want a binding factor that creates mods either up or downstream of binding site. 
# New parameter 'offset.method' to accept function as argument
upDownFunc <- function(x)  {
  return(sample(c(x, -x), 1))
  
}

bf.spreadRep <- createBindingFactor.layer_region("spreadRep", patternLength=150, mismatch.rate=0, 
                                                 profile.layers = "PRC", profile.marks = 1, 
                                                 mod.layers = "H3K27me3", mod.marks=1, 
                                                 stateWidth=200,offset=350, 
                                                 offset.method=upDownFunc)


```





## Run Layer binding

```{r run-binding, echo=TRUE}


n.waves <- 10
n.iters <- 2000

XFS <- list(bf.CpGisland =bf.CpGisland , bf.PRC.1=bf.PRC, bf.spreadRep.1=bf.spreadRep, 
              bf.PRC.2=bf.PRC, bf.spreadRep.2=bf.spreadRep)

XFS.10 <- rep(XFS, times=n.waves)
abundances <- 2000  # runLayerBinding will convert this to a vector as long as XFS.10 so each has 2000 "molecules"

modLayer <- runLayerBinding( layerList=layerList.X, factorSet = XFS.10,
                             bf.abundances =abundances)

# inspect the result
modLayer

```



## Visualise the results
```{r }
#TODO
```



## Score against experimental data
```{r }
#TODO
```
