---
title: "MouseX-inacivation"
author: "Dave T. Gerrard"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MouseX-inacivation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction


A real model of spreading of several factors across the X chromosome.


Requires mouse BSgenome.Mmusculus.UCSC.mm9 which can be obtained from Bioconductor

  if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

  BiocManager::install("BSgenome.Mmusculus.UCSC.mm9")


## Load prerequisites


If  _BSgenome.Mmusculus.UCSC.mm9_ and _GenomicLayers_ are installed, the following should not give errors.

```{r load-libs, echo=TRUE, eval =F}
library(BSgenome.Mmusculus.UCSC.mm9)
library(GenomicLayers)
```




## Set up chromosome and layers
```{r loadGenome, eval =F}
#OUTPUTDIR <- "GL_out/results/mus_X_inactivation/"
GENOME <- 'mm9'

# load X chromosome sequence 
genome <- BSgenome.Mmusculus.UCSC.mm9 

seqinfo(genome)
seqnames(genome)
#sequences_to_keep <- paste0("chr", c(1:19, "X", "Y"))   # nuclear chroms only
sequences_to_keep <- "chrX"   # nuclear chroms only
genomeNuc <- keepBSgenomeSequences(genome, sequences_to_keep)
seqinfo(genomeNuc)



# set up a layerSet on chrom X.

nucLayerSet.X <- createLayerSet.BSgenome( genome=genomeNuc, 
                n.layers= 3,layer.names = c("CpG_island","PRC","H3K27me3"))



```


## Load/Create binding factors
```{r createBfs, eval =F}
#N.B. temporarily setting a fixed patternLength, 
# because don't know how to implement variable patternLength yet.
# current effect is that hits < patternLength are discarded
bf.CpGisland <- createBindingFactor.DNA_regexp("CpGisland", 
                patternString="(CG.{0,20}){9}CG",  
                patternLength=20, mod.layers = "CpG_island", 
                mod.marks=1, stateWidth=200)

bf.EZH2 <- createBindingFactor.layer_region("EZH2",  patternLength=100, mismatch.rate=0, 
                  profile.layers = "CpG_island", profile.marks = 1,  
                  mod.layers = "PRC", mod.marks=1, stateWidth=150)
bf.EZH2$profile$LAYER.0 <- NULL

bf.PRC <- createBindingFactor.layer_region("PRC",  patternLength=100, mismatch.rate=0, 
                  profile.layers = "PRC", profile.marks = 1,  
                  mod.layers = "H3K27me3", mod.marks=1, stateWidth=150)
bf.PRC$profile$LAYER.0 <- NULL

# want a binding factor that creates mods either up or downstream of binding site. 
# New parameter 'offset.method' to accept function as argument
upDownFunc <- function(x)  {
  return(sample(c(x, -x), 1))
  
}

bf.spreadRep <- createBindingFactor.layer_region("spreadRep", patternLength=100,
                  mismatch.rate=0, 
                  profile.layers = "H3K27me3", profile.marks = 1, 
                  mod.layers = "PRC", mod.marks=1, 
                  stateWidth=150,offset=350, 
                  offset.method=upDownFunc)
bf.spreadRep$profile$LAYER.0 <- NULL

bf.spreadRep2 <- createBindingFactor.layer_region("spreadRep2", patternLength=100,
                  mismatch.rate=0, 
                  profile.layers = "H3K27me3", profile.marks = 1, 
                  mod.layers = "PRC", mod.marks=1, 
                  stateWidth=150,offset=500, 
                  offset.method=upDownFunc)
bf.spreadRep2$profile$LAYER.0 <- NULL

```





## Run Layer binding

```{r run-binding, echo=TRUE, eval =F}





XFS <- list(bf.CpGisland =bf.CpGisland ,bf.EZH2=bf.EZH2, bf.PRC=bf.PRC, 
            bf.spreadRep=bf.spreadRep , bf.spreadRep2=bf.spreadRep2 )


abundances <- c(500,5000,5000, 5000,2000)   #

n.iters <- 20
#n.iters <- 2000  #  
                
for (i in 1: n.iters) {
  if(i%%10 == 0)  {
    
    print(i)
  } else {
    #cat(".") 
  }
  nucLayerSet.X <- runLayerBinding.BSgenome( layerList=nucLayerSet.X, 
                                             factorSet = XFS, bf.abundances =abundances, 
                                             collect.stats=T, keep.stats = T)
  
}
# inspect the result
nucLayerSet.X$history

### saturates CpG islands within about 10 iterations but the spread rate slows 
### to cover 2.5Mb (166Mb chrX) after 50 iterations. 7Mb after 200. 20Mb after 2000
### will take a lot more to cover the whole chromosome. 

```



## Visualise the results
```{r , eval =F}
#TODO
plot(nucLayerSet.X$history[,"nBlocks.CpG_island"])
plot(nucLayerSet.X$history[,"Coverage.CpG_island"])
plot(nucLayerSet.X$history[,"nBlocks.PRC"])
plot(nucLayerSet.X$history[,"nBlocks.H3K27me3"])
plot(nucLayerSet.X$history[,"Coverage.PRC"])
plot(nucLayerSet.X$history[,"Coverage.H3K27me3"])
plot(nucLayerSet.X$history[,"time"])
```



## Score against experimental data
```{r , eval =F}
#TODO
```
